What makes listeners infer teaching intentions?

### 2022-02-01 paper/manuscript.Rmd
update results (experiment-1)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

However, the Shapiro-Wilk normality test revealed that the distribution of IOIs for dynamics recordings was not normally distributed. Therefore, a Spearman's rank correlation coefficient was additionally used to assess the relationship between performance tempi of dynamics recordings and participants' judgments. The result failed to reach significance (*r* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_ioi_dyn_spearman$estimate))`, *p* `r if(cor_ioi_dyn_spearman$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_ioi_dyn_spearman$p.value))}`), suggesting that we need to be careful about the interpretation of the relationship between performance tempi of dynamics recordings and participants' judgments as teaching.

@@

@@

@@

@@

Experiment 1 investigated which features of piano performance made musicians (in the role of potential learners) infer teaching intentions. The results demonstrated that performances with slower tempo were more likely to be judged as teaching by musicians regardless which expressive technique was implemented in the piece. For articulation recordings, performances with longer legato and shorter staccato were tended to be judged as teaching. For dynamics recordings, performances with louder forte were more likely to be judged as teaching whereas there was no relationship between softer sound (i.e., piano) and participants' judgements as teaching. However, performances with larger contrasts between forte and piano for both directions (i.e., from forte to piano, from piano to forte) were more likely to be judged as teaching. This result may suggest that dynamics contrast might be reliably used to infer teaching intentions, rather than absolute dynamics values themselves. Moreover, multiple regression analyses implied that tempo feature is the strongest predictor of participants' judgments as teaching in general whereas there were specific predictors depending on which expressive technique was implemented in the piece. These performance features were overall consistent with what expert pianists did in our previous experiments for teaching purposes. Therefore, our findings suggest that musicians may be able to successfully infer teaching intentions of expert pianists, only by listening to recorded performances.

@@

```{r results-2, cache = TRUE, include = FALSE}
source(here("experiment-2/analysis", "analysis.R"), chdir = TRUE)
```

>>>>

However, the Shapiro-Wilk normality test revealed that the distribution of IOIs for dynamics recordings was not normally distributed (*p* `r if(ioi_dyn_norm_performance$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm_performance$p.value))}`). Therefore, a Spearman's rank correlation coefficient was additionally used to assess the relationship between performance tempi of dynamics recordings and participants' judgments. The result failed to reach significance (*r* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_ioi_dyn_spearman$estimate))`, *p* `r if(cor_ioi_dyn_spearman$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', cor_ioi_dyn_spearman$p.value))}`), suggesting that we need to be careful about the interpretation of the relationship between performance tempi of dynamics recordings and participants' judgments as teaching.

@@

The Shapiro-Wilk normality test revealed that the distribution of KOT for dynamics recordings was not normally distributed (Forte; *p* `r if(kot_for_norm_performance$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_for_norm_performance$p.value))}`, Piano; *p* `r if(kot_pia_norm_performance$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_pia_norm_performance$p.value))}`). Spearman's rank correlation coefficients also showed that there was no significant relationship between KOT values and participants' judgments as teaching (Forte; *r* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_for_spearman$estimate))`, *p* `r if(cor_kot_for_spearman$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_kot_for_spearman$p.value))}`, Piano; *r* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_kot_pia_spearman$estimate))`, *p* `r if(cor_kot_pia_spearman$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_kot_pia_spearman$p.value))}`).

@@

The Shapiro-Wilk normality test revealed that the distribution of KOT for articulation recordings for legato was not normally distributed (;*p* `r if(vel_leg_norm_performance$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_leg_norm_performance$p.value))}`). A Spearman's rank correlation coefficients also showed that there was no significant relationship between KOT values for legato and participants' judgments as teaching (*r* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_leg_spearman$estimate))`, *p* `r if(cor_vel_leg_spearman$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_vel_leg_spearman$p.value))}`).

@@

The Shapiro-Wilk normality test revealed that the distribution of KV difference for articulation recordings for transition points from legato to staccato was not normally distributed (;*p* `r if(vel_diff_ltos_norm_performance$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_ltos_norm_performance$p.value))}`). A Spearman's rank correlation coefficients also showed that there was no significant relationship between KOT values for legato and participants' judgments as teaching (*r* = `r sub("^(-?)0.", "\\1.", sprintf('%.2f', cor_vel_diff_ltos_spearman$estimate))`, *p* `r if(cor_vel_diff_ltos_spearman$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', cor_vel_diff_ltos_spearman$p.value))}`).

@@

Experiment 1 investigated which features of piano performance made musicians (in the role of potential learners) infer teaching intentions. The results demonstrated that performances with slower tempo were more likely to be judged as teaching by musicians regardless which expressive technique was implemented in the piece. For articulation recordings, performances with longer legato and shorter staccato were tended to be judged as teaching. For dynamics recordings, performances with louder forte were more likely to be judged as teaching whereas there was no relationship between softer sound (i.e., piano) and participants' judgments as teaching. Importantly, performances with larger contrasts between forte and piano for both directions (i.e., from forte to piano, from piano to forte) were more likely to be judged as teaching. This result may suggest that dynamics contrast might be reliably used to infer teaching intentions, rather than absolute dynamics values themselves. Moreover, multiple regression analyses implied that tempo feature is the strongest predictor of participants' judgments as teaching in general whereas there were specific predictors depending on which expressive technique was implemented in the piece. These performance features were overall consistent with what expert pianists did in our previous experiments for teaching purposes. Therefore, our findings suggest that musicians may rely on generic pedagogical behaviour (e.g., slower demonstration, exaggeration) to infer teaching intentions of expert pianists, only by listening to recorded performances.

@@

```{r results-2, cache = FALSE, include = FALSE}
source(here("experiment-2/analysis", "analysis.R"), chdir = TRUE)
```

### 2022-02-01 paper/manuscript.tex
update results (experiment-1)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

The experiment was programmed in Python 3.8.2 using the PsychoPy library (2020.2.4; \url{https://www.psychopy.org/}) on a Mac Book Pro with Mac OS X Catalina 10.15.6. Stimuli were played using the Mido library (1.2.9; \url{https://mido.readthedocs.io/en/latest/}) on a Max/MSP patcher (8.1.7; \url{https://cycling74.com/products/max}). During the experiment, participants listened to the stimuli via headphones (Audio-Technica ATH-M50X).

@@

For the current experiment, 96 performances were chosen from the valid performances in Tominaga et al.~(\emph{submitted}). We randomly sampled 24 articulation performances and 24 dynamics performances from the teaching condition as well as 24 articulation performances and 24 dynamics performances from the performing condition. It is important to note that the performances from the teaching condition did not necessarily exhibit specific features of teaching performance that we found in our previous experiments since we randomly sampled the recordings.

@@

First, we investigated data by looking at correlations between performance features (i.e., IOIs, KOT, KV, KV-Diff) and participants' judgments as ``teaching'' (i.e., what percentage of participants responded as ``yes (teaching)'') for each stimulus. Second, we run multiple regression and entered all four performance features to examine which predictor significantly contributed to participants' judgments as teaching. Since articulation and dynamics were comprised of two opposite directional values (i.e., legato vs.~staccato, forte vs.~piano), we entered each subcomponent (i.e., legato, staccato, forte, piano) separately to represent either an articulation or dynamics feature of the performance for a regression model.

@@

However, the Shapiro-Wilk normality test revealed that the distribution of IOIs for dynamics recordings was not normally distributed. Therefore, a Spearman's rank correlation coefficient was additionally used to assess the relationship between performance tempi of dynamics recordings and participants' judgments. The result failed to reach significance (\emph{r} = .27, \emph{p} = 0.066), suggesting that we need to be careful about the interpretation of the relationship between performance tempi of dynamics recordings and participants' judgments as teaching.

@@

@@

@@

>>>>

The experiment was programmed in Python 3.8.2 using the PsychoPy Python library (2020.2.4; \url{https://www.psychopy.org/}) on a Mac Book Pro with Mac OS X Catalina 10.15.6. Stimuli were played using the Mido Python library (1.2.9; \url{https://mido.readthedocs.io/en/latest/}) on a Max/MSP patcher (8.1.7; \url{https://cycling74.com/products/max}). During the experiment, participants listened to the stimuli via headphones (Audio-Technica ATH-M50X).

@@

For the current experiment, 96 performances were chosen from the valid performances in Tominaga et al.~(\emph{submitted}). We randomly sampled 24 articulation performances and 24 dynamics performances from the teaching condition as well as 24 articulation performances and 24 dynamics performances from the performing condition. It is important to note that the performances from the teaching condition did not necessarily exhibit specific features of teaching that we found in our previous experiments (e.g., exaggeration) since we randomly sampled the performances.

@@

First, we investigated data by looking at correlations between performance features (i.e., IOIs, KOT, KV, KV-Diff) and participants' judgments as ``teaching'' (i.e., what percentage of participants responded as ``yes'') for each stimulus. Second, we run multiple regression and entered all four performance features to examine which predictor significantly contributed to participants' judgments as teaching. Since articulation and dynamics were comprised of two opposite directional values (i.e., legato vs.~staccato, forte vs.~piano), we entered each subcomponent (i.e., legato, staccato, forte, piano) separately to represent either an articulation or dynamics feature of the performance for a regression model.

@@

However, the Shapiro-Wilk normality test revealed that the distribution of IOIs for dynamics recordings was not normally distributed (\emph{p} = 0.037). Therefore, a Spearman's rank correlation coefficient was additionally used to assess the relationship between performance tempi of dynamics recordings and participants' judgments. The result failed to reach significance (\emph{r} = .27, \emph{p} = 0.066), suggesting that we need to be careful about the interpretation of the relationship between performance tempi of dynamics recordings and participants' judgments as teaching.

@@

The Shapiro-Wilk normality test revealed that the distribution of KOT for dynamics recordings was not normally distributed (Forte; \emph{p} \textless{} 0.001, Piano; \emph{p} \textless{} 0.001). Spearman's rank correlation coefficients also showed that there was no significant relationship between KOT values and participants' judgments as teaching (Forte; \emph{r} = .15, \emph{p} = 0.31, Piano; \emph{r} = -.12, \emph{p} = 0.43).

@@

The Shapiro-Wilk normality test revealed that the distribution of KOT for articulation recordings for legato was not normally distributed (;\emph{p} = 0.002). A Spearman's rank correlation coefficients also showed that there was no significant relationship between KOT values for legato and participants' judgments as teaching (\emph{r} = .03, \emph{p} = 0.84).

@@

The Shapiro-Wilk normality test revealed that the distribution of KV difference for articulation recordings for transition points from legato to staccato was not normally distributed (;\emph{p} = 0.007). A Spearman's rank correlation coefficients also showed that there was no significant relationship between KOT values for legato and participants' judgments as teaching (\emph{r} = -.23, \emph{p} = 0.12).
